---
title: "Reannotating the DMRs"
author: "Andrew Y.F. Li Yim"
date: "June/July 2017"
output:
  pdf_document: 
    highlight: tango
    number_sections: yes
    toc: yes
---

Step 1.1 Reannotating the DMRs
===============================
In this workbook we aim to reannotate the DMRs by associating them to genes on the basis of their overlap with known promoters or enhancers (regulatory features). For the promoters we will use the Ensembl Regulatory Build (v89), while the enhancers will be based on promoter capture Hi-C (pcHi-C) overlayed on top of the Ensembl Regulatory Build (v89). As promoter/enhancer activity is celltype specific, we will confine our reannotation to the celltypes that are were present in both promoter and enhancer datasets, namely: Monocytes, Macrophages 0, Macrophages 1, Macrophages 2, Neutrophils, Erythroblasts, CD4 T-cells, and CD8 T-cells.

# Preparation
```{r Filepaths}

```


```{r Importing the DMRs}
DMR_path <- "../../0_Original_project/data/analysis/WGBS/07differentialMethylation/"

HvL_dmrs <- read.csv(paste0(DMR_path, "HL.dmr.csv"), stringsAsFactors = F)
HvN_dmrs <- read.csv(paste0(DMR_path, "HN.dmr.csv"), stringsAsFactors = F)
LvN_dmrs <- read.csv(paste0(DMR_path, "LN.dmr.csv"), stringsAsFactors = F)

require(GenomicRanges)
HvL_dmrs_gr <- makeGRangesFromDataFrame(df = HvL_dmrs, keep.extra.columns = T, seqnames.field = "chr", start.field = "start", end.field = "end")
HvN_dmrs_gr <- makeGRangesFromDataFrame(df = HvN_dmrs, keep.extra.columns = T, seqnames.field = "chr", start.field = "start", end.field = "end")
LvN_dmrs_gr <- makeGRangesFromDataFrame(df = LvN_dmrs, keep.extra.columns = T, seqnames.field = "chr", start.field = "start", end.field = "end")
```

```{r Genomic reannotation resources}
require(rtracklayer)
require(ensembldb)
require(AnnotationDbi)
require(GenomicFeatures)
require(org.Hs.eg.db)
require(ChIPseeker)

hg19tohg38 <- import.chain("../../0_Original_project/data/analysis/WGBS/00correlations/hg19ToHg38.over.chain")
anno_folder <- "/home/al580162/data/ensembl_annotation/"
anno_GRCh37 <- file.path(anno_folder, "Homo_sapiens.GRCh37.75.txdb.db")
anno_GRCh38 <- file.path(anno_folder, "Homo_sapiens.GRCh38.89.txdb.db")

if(file.exists(anno_GRCh38)){
	txdb_GRCh38 <- loadDb(anno_GRCh38)
} else{
	txdb_GRCh38 <- makeTxDbFromGFF(file.path(anno_folder, "Homo_sapiens.GRCh38.89.gtf"), taxonomyId = 9606)
	txdb_GRCh38 <- renameSeqlevels(txdb_GRCh38, gsub("^([0-9XY]+)$", "chr\\1", seqlevels(txdb_GRCh38)))
	saveDb(x = txdb_GRCh38, file = anno_GRCh38)
}
if(file.exists(anno_GRCh37)){
	txdb_GRCh37 <- loadDb(anno_GRCh37)
} else{
	txdb_GRCh37 <- makeTxDbFromGFF(file.path(anno_folder, "Homo_sapiens.GRCh37.75.gtf"), taxonomyId = 9606)
	txdb_GRCh37 <- renameSeqlevels(txdb_GRCh37, gsub("^([0-9XY]+)$", "chr\\1", seqlevels(txdb_GRCh37)))
	saveDb(x = txdb_GRCh37, file = anno_GRCh37)
}
```


# Promoters
To find the overlap between the DMRs and the promoters, we will utilize ChIPSeeker (Yu et al. 2015). One problem is that for some blood cells, there is only venous blood (VB) or cord blood (CB). A moderate solution is to download both for each dataset and merge the active regions from both. 

## Importing the promoters
```{r Importing and processing the Ensembl Regulatory Build v89}
promoter_data <- "../Data/EnsemblRegulatoryBuild/v89"
ensreg_promoters <- read.csv(file.path(promoter_data, "GRCh38_promoters_biomart.tsv"), stringsAsFactors = F, header = T, sep = "\t")
ensreg_promoters_df <- data.frame(chr = ensreg_promoters[, 1],
				  start = ensreg_promoters[, 2],
				  end = ensreg_promoters[, 3],
				  feature_type = ensreg_promoters[, 4],
				  feature_status = ensreg_promoters[, 6],
				  celltype = gsub("(^.+)\\(((C|V)B)\\)$", "\\1_\\2", ensreg_promoters[, 5]))
ensreg_promoters_df <- unique(ensreg_promoters_df)

ensreg_promoters_df$celltype <- gsub("^CD14\\+.+(_(C|V)B)$", "Mon\\1", ensreg_promoters_df$celltype)
ensreg_promoters_df$celltype <- gsub("^CD4\\+.+(_(C|V)B)$", "CD4\\1", ensreg_promoters_df$celltype)
ensreg_promoters_df$celltype <- gsub("^CD8\\+.+(_(C|V)B)$", "CD8\\1", ensreg_promoters_df$celltype)
ensreg_promoters_df$celltype <- gsub("^M0.+(_(C|V)B)$", "Mac0\\1", ensreg_promoters_df$celltype)
ensreg_promoters_df$celltype <- gsub("^M1.+(_(C|V)B)$", "Mac1\\1", ensreg_promoters_df$celltype)
ensreg_promoters_df$celltype <- gsub("^M2.+(_(C|V)B)$", "Mac2\\1", ensreg_promoters_df$celltype)
ensreg_promoters_df$celltype <- gsub("erythro.+(_(C|V)B)$", "Ery\\1", ensreg_promoters_df$celltype)
ensreg_promoters_df$celltype <- gsub("neutro.+(_(C|V)B)$", "Neu\\1", ensreg_promoters_df$celltype)
ensreg_promoters_df <- unique(ensreg_promoters_df)

ensreg_promoters_df <- aggregate(celltype~., ensreg_promoters_df, FUN = toString)
ensreg_promoters_df$celltype <- gsub(", ", ";", ensreg_promoters_df$celltype)

ensreg_promoters_gr <- keepStandardChromosomes(makeGRangesFromDataFrame(df = ensreg_promoters_df, keep.extra.columns = T), pruning.mode = "coarse")
ensreg_promoters_gr <- ensreg_promoters_gr[ensreg_promoters_gr$feature_type == "Promoter", ]

ensreg_promoters_anno <- as.GRanges(annotatePeak(ensreg_promoters_gr, tssRegion = c(-2500, 250), TxDb = txdb_GRCh38, annoDb = "org.Hs.eg.db"))
write.csv(as.data.frame(ensreg_promoters_anno), file.path(promoter_data, "GRCh38_promoters_blood.csv"))
ensreg_actprom_anno <- ensreg_promoters_anno[which(ensreg_promoters_anno$feature_status == "ACTIVE"), ]
write.csv(as.data.frame(ensreg_actprom_anno), file.path(promoter_data, "GRCh38_actprom_blood.csv"))

```

## Overlapping the DMRs with the promoters
```{r Overlapping the DMRs with the newly annotated promoters}
dm_promoter_overlap <- function(dmrs, promoters){
	dmrs <- keepStandardChromosomes(dmrs, pruning.mode = "coarse")
	promoters <- keepStandardChromosomes(promoters, pruning.mode = "coarse")

	overlapping_ind <- findOverlaps(promoters, dmrs)
	if(length(overlapping_ind) != 0){

		dmrs_df <- data.frame(dmrs)[subjectHits(overlapping_ind), c("seqnames", "start", "end", "n", "invdensity", "areaStat", "maxStat", "meanDiff", "group1.mean", "group2.mean", "tstat.sd", "direction")]
		colnames(dmrs_df) <- c("dmrChr", "dmrStart", "dmrEnd", "nCpGs", "invdensity", "areaStat", "maxStat", "meanDiff", "group1.mean", "group2.mean", "tstat.sd", "direction")

		promoters_df <- data.frame(promoters)[queryHits(overlapping_ind), c("seqnames", "start", "end", "feature_type", "feature_status", "celltype", "distanceToTSS", "geneChr", "geneStart", "geneEnd", "geneId", "SYMBOL")]
		colnames(promoters_df) <- c("promChr", "promStart", "promEnd", "feature_type", "feature_status", "celltype",  "promDistToTSS", "geneChr", "geneStart", "geneEnd", "geneENS", "geneSYMBOL")

		dm_promoters <- cbind(dmrs_df, promoters_df)

		dm_promoters$dmrChr <- gsub("chr", "", dm_promoters$dmrChr)
		dm_promoters$promChr <- gsub("chr", "", dm_promoters$promChr)
		dm_promoters$geneChr <- gsub("chr", "", dm_promoters$geneChr)

		dm_promoters <- aggregate(celltype~., dm_promoters, FUN = toString)
		dm_promoters$celltype <- gsub(", *", ";", dm_promoters$celltype)
		return(dm_promoters)
	}
	else{
		cat("No overlaps were found\n")
	}
}

HvL_dm_prom <- dm_promoter_overlap(HvL_dmrs_gr, ensreg_promoters_anno)
HvN_dm_prom <- dm_promoter_overlap(HvN_dmrs_gr, ensreg_promoters_anno)
LvN_dm_prom <- dm_promoter_overlap(LvN_dmrs_gr, ensreg_promoters_anno)
write.csv(HvL_dm_prom, "HvL_dmpromoters.csv")
write.csv(HvN_dm_prom, "HvN_dmpromoters.csv")
write.csv(LvN_dm_prom, "LvN_dmpromoters.csv")

HvL_dm_actprom <- HvL_dm_prom[HvL_dm_prom$feature_status == "ACTIVE",]
HvN_dm_actprom <- HvN_dm_prom[HvN_dm_prom$feature_status == "ACTIVE",]
LvN_dm_actprom <- LvN_dm_prom[LvN_dm_prom$feature_status == "ACTIVE",]
write.csv(HvL_dm_actprom, "HvL_dm_active_promoters.csv")
write.csv(HvN_dm_actprom, "HvN_dm_active_promoters.csv")
write.csv(LvN_dm_actprom, "LvN_dm_active_promoters.csv")
```


# Enhancers
To find the overlap between the DMRs and the promoters, we will again utilize ChIPSeeker (Yu et al. 2015). The pHi-C data is obtained from the Javierre et al. 2016 article, in particular three files were provided by Javierre et al. 2016, namely:
* PCHiC_peak_matrix_cutoff5.tsv: Contains all the promoter-chromatin interactions with a CHiCAGO score above 5.
* PCHiC_vs_rCHiC_peak_matrix.tsv: Promoter-chromatin interactions that were technically validated by performing the inverse capture.
* ActivePromoterEnhancerLinks.tsv: Promoter-chromatin interactions that were also found in the Ensembl Regulatory Build and BLUEPRINT data.
We will use the last dataset in our analyses.

## Importing the enhancers
```{r Importing and processing the ActivePromoterEnhancerLinks.tsv}
ActiveProm <- read.csv("../Data/Javierre2016/ActivePromoterEnhancerLinks.tsv", sep = "\t")
#ActiveProm$baitChr <- gsub("chr", "", ActiveProm$baitChr)
#ActiveProm$oeChr <- gsub("chr", "", ActiveProm$oeChr)

celltypes_assayed <- unique(unlist(strsplit(x = as.character(unique(ActiveProm$cellType.s.)), split = ",")))

ActivePromOE_gr <- makeGRangesFromDataFrame(df = ActiveProm, keep.extra.columns = T, seqnames.field = "oeChr", start.field = "oeSt", end.field = "oeEnd")
seqlevelsStyle(ActivePromOE_gr) <- "NCBI"
ActivePromOE_anno <- as.GRanges(annotatePeak(ActivePromOE_gr, tssRegion = c(-2500, 250), TxDb = txdb_GRCh37, annoDb = "org.Hs.eg.db"))
seqlevelsStyle(ActivePromOE_anno) <- "UCSC"
ActivePromOE_anno <- unlist(liftOver(ActivePromOE_anno, hg19tohg38)) 
seqlevelsStyle(ActivePromOE_anno) <- "NCBI"
ActivePromOE_anno_df <- as.data.frame(ActivePromOE_anno)
ActivePromOE_anno_df_sub <- ActivePromOE_anno_df[,c("seqnames", "start", "end", "baitChr", "baitSt", "baitEnd", "annotation", "distanceToTSS", "geneId", "SYMBOL", "cellType.s.")]
colnames(ActivePromOE_anno_df_sub) <- c("oeChr_GRCh38", "oeStart_GRCh38", "oeEnd_GRCh38", "baitChr_GRCh37", "baitStart_GRCh37", "baitEnd_GRCh37", "oeAnno", "oeDistTSS", "oeENS", "oeSYMBOL", "celltype")

ActivePromBait_gr <- makeGRangesFromDataFrame(df = ActivePromOE_anno_df_sub, keep.extra.columns = T, seqnames.field = "baitChr_GRCh37", start.field = "baitStart_GRCh37", end.field = "baitEnd_GRCh37")
seqlevelsStyle(ActivePromBait_gr) <- "NCBI"
ActivePromBait_anno <- as.GRanges(annotatePeak(ActivePromBait_gr, tssRegion = c(-2500, 250), TxDb = txdb_GRCh37, annoDb = "org.Hs.eg.db"))
seqlevelsStyle(ActivePromBait_anno) <- "UCSC"
ActivePromBait_anno <- unlist(liftOver(ActivePromBait_anno, hg19tohg38))
seqlevelsStyle(ActivePromBait_anno) <- "NCBI"
ActivePromBait_anno_df <- as.data.frame(ActivePromBait_anno)
ActivePromBait_anno_df_sub <- ActivePromBait_anno_df[,c("seqnames", "start", "end", "annotation", "distanceToTSS", "geneId", "SYMBOL","oeChr_GRCh38", "oeStart_GRCh38", "oeEnd_GRCh38", "oeAnno", "oeDistTSS", "oeENS", "oeSYMBOL", "celltype")]
colnames(ActivePromBait_anno_df_sub) <- c("baitChr", "baitStart", "baitEnd", "baitAnno", "baitDistTSS", "baitENS", "baitSYMBOL", "oeChr", "oeStart", "oeEnd", "oeAnno", "oeDistTSS", "oeENS", "oeSYMBOL", "celltype")

write.csv(ActivePromBait_anno_df_sub, "ActiveProm_anno_GRCh38.csv")

ActiveProm_GRCh38_OE <- makeGRangesFromDataFrame(df = ActivePromBait_anno_df_sub, keep.extra.columns = T, seqnames.field = "oeChr", start.field = "oeStart", end.field = "oeEnd")
```

Some enhancers appear to overlap with other enhancers, it is unclear why these hits were not merged.
```{r Merge overlapping enhancers NOT FINISHED}
ActiveProm_GRCh38_OE_split <- split(ActiveProm_GRCh38_OE, ActiveProm_GRCh38_OE$oeENS)
ActiveProm_GRCh38_OE_red <- reduce(ActiveProm_GRCh38_OE)
overlapping_enhancers <- data.frame(findOverlaps(query = ActiveProm_GRCh38_OE, subject = ActiveProm_GRCh38_OE_red))
overlapping_enhancers_list <- split(overlapping_enhancers, overlapping_enhancers$subjectHits)
lapply(overlapping_enhancers_list, function(entry){
	split(entry, entry$oeENS)
})

```
## Overlapping the DMRs with the enhancers
```{r Overlapping the DMRs with the newly annotated enhancers}
dm_enhancer_overlap <- function(dmrs, enhancers){
	dmrs <- keepStandardChromosomes(dmrs, pruning.mode = "coarse")
	enhancers <- keepStandardChromosomes(enhancers, pruning.mode = "coarse")

	overlapping_ind <- findOverlaps(enhancers, dmrs)
	dmrs_df <- data.frame(dmrs)[subjectHits(overlapping_ind), c("seqnames", "start", "end", "n", "invdensity", "areaStat", "maxStat", "meanDiff", "group1.mean", "group2.mean", "tstat.sd", "direction")]
	colnames(dmrs_df) <- c("dmrChr", "dmrStart", "dmrEnd", "nCpGs", "invdensity", "areaStat", "maxStat", "meanDiff", "group1.mean", "group2.mean", "tstat.sd", "direction")

	enhancers_df <- data.frame(enhancers)[queryHits(overlapping_ind), c("seqnames", "start", "end", "oeAnno", "oeDistTSS", "oeENS", "oeSYMBOL", "baitChr", "baitStart", "baitEnd", "baitAnno", "baitDistTSS", "baitENS", "baitSYMBOL", "celltype")]
	colnames(enhancers_df) <- c("oeChr", "oeStart", "oeEnd", "oeAnno", "oeDistTSS", "oeENS", "oeSYMBOL", "baitChr", "baitStart", "baitEnd", "baitAnno", "baitDistTSS", "baitENS", "baitSYMBOL", "celltype")

	dm_enhancers <- cbind(dmrs_df, enhancers_df)

	dm_enhancers$dmrChr <- gsub("chr", "", dm_enhancers$dmrChr)
	dm_enhancers$oeChr <- gsub("chr", "", dm_enhancers$oeChr)
	dm_enhancers$baitChr <- gsub("chr", "", dm_enhancers$baitChr)

	dm_enhancers <- aggregate(celltype~., dm_enhancers, FUN = toString)
	dm_enhancers$celltype <- gsub(", *", ";", dm_enhancers$celltype)

	return(dm_enhancers)
}

HvL_dmen <- dm_enhancer_overlap(HvL_dmrs_gr, ActiveProm_GRCh38_OE) 
HvN_dmen <- dm_enhancer_overlap(HvN_dmrs_gr, ActiveProm_GRCh38_OE)
LvN_dmen <- dm_enhancer_overlap(LvN_dmrs_gr, ActiveProm_GRCh38_OE)
write.csv(HvL_dmen, "HvL_dmenhancers.csv")
write.csv(HvN_dmen, "HvN_dmenhancers.csv")
write.csv(LvN_dmen, "LvN_dmenhancers.csv")
```

# Statistics 
## Cellular overrepresentation
Cellular oVerrepresentation analysis: Can we find more DMRs for a particular cell type? On a second thought, this does not make a lot of sense, what would it mean to find more DMRs in the promoters/enhancers for a particular celltype. The interpretability is lost.

Certain functions to be used throughout this notebook
```{r Functions}
groupwise_fishtest <- function(df){
	stats <- matrix(ncol = 5, nrow = ncol(df))
	colnames(stats) <- c("pval", "padj", "lower_CI95", "upper_CI95", "OR")
	rownames(stats) <- colnames(df)
	for(i in 1:ncol(df)){
		fish_res <- fisher.test(data.frame(case = df[,i], control = rowSums(df[, -i])))
		stats[i, "pval"] <- fish_res$p.value
		stats[i, "lower_CI95"] <- fish_res$conf.int[1]
		stats[i, "upper_CI95"] <- fish_res$conf.int[2]
		stats[i, "OR"] <- fish_res$estimate
	}
	stats[,"padj"] <- p.adjust(stats[,"pval"])	
	return(stats)
}
```
```{r DMpromoter statistics}
# Separate VB/CB
HvL_dmprom_cells <- table(unlist(strsplit(as.character(HvL_dmprom$celltype), ";")))
HvN_dmprom_cells <- table(unlist(strsplit(as.character(HvN_dmprom$celltype), ";")))
LvN_dmprom_cells <- table(unlist(strsplit(as.character(LvN_dmprom$celltype), ";")))
BG_prom_cells <- table(unlist(strsplit(as.character(ensreg_actprom_anno$celltype), ";")))

HvLvBG_prom <- groupwise_fishtest(rbind(HvL_dmprom_cells, BG_prom_cells))
HvNvBG_prom <- groupwise_fishtest(rbind(HvN_dmprom_cells, BG_prom_cells))
write.csv(HvLvBG_prom, "HvLvBG_prom_fisher.csv") 
write.csv(HvNvBG_prom, "HvNvBG_prom_fisher.csv")

#Merged VB/CB
HvL_dmprom_cells_merged <- table(unlist(lapply(strsplit(gsub("_(C|V)B", "", HvL_dmprom$celltype), ";"), unique)))
HvN_dmprom_cells_merged <- table(unlist(lapply(strsplit(gsub("_(C|V)B", "", HvN_dmprom$celltype), ";"), unique)))
BG_prom_cells_merged <- table(unlist(lapply(strsplit(gsub("_(C|V)B", "", ensreg_actprom_anno$celltype), ";"), unique)))

HvLvBG_prom_merged <- groupwise_fishtest(rbind(HvL_dmprom_cells_merged, BG_prom_cells_merged))
HvNvBG_prom_merged <- groupwise_fishtest(rbind(HvN_dmprom_cells_merged, BG_prom_cells_merged))
write.csv(HvLvBG_prom_merged, "HvLvBG_prom_merged_fisher.csv")
write.csv(HvNvBG_prom_merged, "HvNvBG_prom_merged_fisher.csv")

```
```{r DMpromoter statistics}
HvL_dmen_cells <- table(unlist(strsplit(as.character(HvL_dmen$celltype), ";")))
HvN_dmen_cells <- table(unlist(strsplit(as.character(HvN_dmen$celltype), ";")))
LvN_dmen_cells <- table(unlist(strsplit(as.character(LvN_dmen$celltype), ";")))
LvN_dmen_cells <- c(0, 0, 0, 2, 2, 0, 0, 0, 0) # Bit of a hack, the problem is that the previous function would only return a"2,2"
BG_dmen_cells <- table(unlist(strsplit(as.character(ActiveProm_GRCh38_OE$celltype), ",")))

HvLvBG_perc <- HvL_dmen_cells/BG_dmen_cells*100
HvNvBG_perc <- HvN_dmen_cells/BG_dmen_cells*100
LvNvBG_perc <- LvN_dmen_cells/BG_dmen_cells*100

#Hypergeometric test
HvLvBG_dmen <- rbind(HvL_dmen_cells, BG_dmen_cells)
HvNvBG_dmen <- rbind(HvN_dmen_cells, BG_dmen_cells)
LvNvBG_dmen <- rbind(LvN_dmen_cells, BG_dmen_cells)

HvLvBG_dmen_fisher <- groupwise_fishtest(HvLvBG_dmen)
HvNvBG_dmen_fisher <- groupwise_fishtest(HvNvBG_dmen)
LvNvBG_dmen_fisher <- groupwise_fishtest(LvNvBG_dmen)

write.csv(HvLvBG_dmen_fisher, "HvLvBG_dmen_fisher.csv")
write.csv(HvNvBG_dmen_fisher, "HvNvBG_dmen_fisher.csv")
write.csv(LvNvBG_dmen_fisher, "LvNvBG_dmen_fisher.csv")
```

## Regulatory region overrepresentation analysis
In total we find 175, 163, and 330  unique DMRs that associate to promoters, enhancers, and regulators (promoters+enhancers) respectively. Note that 175+163 > 330. The discrepancy is because some DMRs are enhancers to one and promoters to another.

Import the CpG coordinates for reshuffling
```{r CpG coordinates}
fit2 <- readRDS(file.path(DMR_path, "fit2.rds"))
CpG_locs <- granges(fit2)
CpG_locs <- keepStandardChromosomes(CpG_locs, pruning.mode = "coarse")
CpG_locs <- split(CpG_locs, seqnames(CpG_locs))

#Function for a single DMR. To me it makes computationally more sense to permute on a per DMR basis rather than a per bootstrap basis. In the case of the latter, for every iteration, each DMR must be checked against the list of regulators.
DMR_shuffler_single <- function(dmr, cpgs, B){
	chro <- paste0(seqnames(dmr))
	cpgs_chro <- as.data.frame(cpgs[[chro]])
	nCpGs <- dmr$nCpGs
	sample_space <- nrow(cpgs_chro)-nCpGs
	indices <- sample(x = 1:sample_space, size = B, replace = T)
	randmrs <- data.frame(chr = cpgs_chro[indices, "seqnames"],
			      start = cpgs_chro[indices, "start"],
			      end = cpgs_chro[indices+nCpGs, "end"])

	return(randmrs)
}

#Wrapper function for all DMRs
require(foreach)
DMRs_shuffler <- function(dmrs, cpgs, B, cores){

	if(cores > 1){
		require(parallel)
		cl <- makeCluster(cores)
		clusterExport(cl, c("cpgs", "B", "DMR_shuffler_single"), envir = environment())
		randmrs_list <- parLapply(cl, dmrs, function(dmr){
			DMR_shuffler_single(dmr, cpgs, B)
		})
		stopCluster(cl)
	} else{
		#First shuffle each DMR B times
		randmrs_list <- lapply(dmrs, function(dmr){
			DMR_shuffler_single(dmr, cpgs, B)		      
		})
	}

	#"Transpose" the list to ensure that each iterations is a list in itself containing the randomized dataframe for that particular DMR
	randmrs_t <- lapply(1:B, function(index){
		do.call(rbind, lapply(randmrs_list, `[`, index, ))
	})

	#Convert the df into a GRanges
	randmrs_gr <- lapply(randmrs_t, makeGRangesFromDataFrame)
	return(randmrs_gr)
}

#Calculate the null distribution
nulldist <- function(randmrs_list_gr, regulators_gr){
	lapply(randmrs_list_gr, function(randmrs){
		length(subsetByOverlaps(randmrs, regulators_gr))
	})
}
```

```{r Generating the null distribution - SEEMS TO ME SOMETHING IS WRONG}

total_dmrs <- rbind(HvL_dmrs[, c(1:3, 7)], HvN_dmrs[, c(1:3, 7)], LvN_dmrs[, c(1:3, 7)])
colnames(total_dmrs) <- c("chr", "start", "end", "nCpGs")
total_dmrs_gr <- unique(makeGRangesFromDataFrame(total_dmrs, keep.extra.columns = T))

rand_dmrs <- DMRs_shuffler(total_dmrs_gr, CpG_locs, 1000, 4)
```


```{r Promoter}
unique_promoter_DMRs <- unique(rbind(HvL_dm_prom[, 1:4], HvN_dm_prom[, 1:4], LvN_dm_prom[, 1:4]))
unique_promoter_DMRs_gr <- makeGRangesFromDataFrame(unique_promoter_DMRs, keep.extra.columns = T, seqnames.field = "dmrChr", start.field = "dmrStart", end.field = "dmrEnd")

nulldist_promdmrs <- unlist(nulldist(rand_dmrs, unique(ensreg_promoters_gr)))

mean(length(unique_promoter_DMRs_gr) <= nulldist_promdmrs)

```

```{r Enhancer}
unique_enhancer_DMRs <- unique(rbind(HvL_dmen[, 1:4], HvN_dmen[, 1:4], LvN_dmen[, 1:4]))
unique_enhancer_DMRs_gr <- makeGRangesFromDataFrame(unique_enhancer_DMRs, keep.extra.columns = T, seqnames.field = "dmrChr", start.field = "dmrStart", end.field = "dmrEnd")

nulldist_enhdmrs <- unlist(nulldist(rand_dmrs, unique(ActiveProm_GRCh38_OE)))

```

```{r Both}
unique_regulator_DMRs <- unique(rbind(unique_promoter_DMRs, unique_enhancer_DMRs))

```


