---
title: "Integrating GWAS"
author: "Andrew Y.F. Li Yim"
date: "August 2017"
output:
  pdf_document:
      highlight: tango
          number_sections: yes
	      toc: yes
---

```{r Filepaths}
data_dir <- file.path("../../data")
output_dir <- file.path("../../output")
eQTM_analyses_dir <- file.path(output_dir, "02_eQTM_analyses")

#Primary analyses
primary_analyses_dir <- file.path(data_dir, "Primary_analyses")

#Methylation analyses
DMR_dir <- file.path(primary_analyses_dir, "data/analysis/WGBS/07differentialMethylation")

#Expression analyses
exprs_dir <- file.path(primary_analyses_dir, "data/analysis/RNAseq/07countReads")
degs_dir <- file.path(primary_analyses_dir, "data/analysis/RNAseq/08differentialExpression")

#Parsed regulatory elements
parsed_promoter_dir <- file.path(output_dir, "01_DMR_annotation/Promoters_parsed")
parsed_enhancer_dir <- file.path(output_dir, "01_DMR_annotation/Enhancers_parsed")

#Semi-annotated DMRs
DMenhancer_dir <- file.path(output_dir, "01_DMR_annotation/DM_enhancers")
DMpromoter_dir <- file.path(output_dir, "01_DMR_annotation/DM_promoters")
```

```{r aDMRs}
require(GenomicRanges)
admr_dir <- file.path(eQTM_analyses_dir, "aDMRs")

admrs <- read.csv(file.path(admr_dir, "aDMRs.csv"), stringsAsFactors = F)[, -1]
admrs_gr <- makeGRangesFromDataFrame(admrs, keep.extra.columns = T)

dadmrs <- read.csv(file.path(admr_dir, "DE_aDMRs.csv"), stringsAsFactors = F)[, -1] 
dadmrs_gr <- makeGRangesFromDataFrame(dadmrs, keep.extra.columns = T)

eqtms <- read.csv(file.path(admr_dir, "eQTMs.csv"), stringsAsFactors = F)[, -1]
eqtms_gr <- makeGRangesFromDataFrame(eqtms, keep.extra.columns = T)

deqtms <- read.csv(file.path(admr_dir, "DeQTMs.csv"), stringsAsFactors = F)[, -1]
deqtms_gr <- makeGRangesFromDataFrame(deqtms, keep.extra.columns = T)
```

```{r Expression}
HvL_degs <- read.csv(file.path(degs_dir, "HL.deg.csv"), stringsAsFactors = F)
HvL_degs <- HvL_degs[order(HvL_degs$pvalue), ]
HvL_degs_sig <- HvL_degs[HvL_degs$padj < 0.05, ]
HvL_degs_sig <- HvL_degs_sig[-unique(which(is.na(HvL_degs_sig), arr.ind = T)[, 1]), ]
HvL_degs_sig$contrast <- "HvL"

HvN_degs <- read.csv(file.path(degs_dir, "HN.deg.csv"), stringsAsFactors = F)
HvN_degs <- HvN_degs[order(HvN_degs$pvalue), ]
HvN_degs_sig <- HvN_degs[HvN_degs$padj < 0.05, ]
HvN_degs_sig <- HvN_degs_sig[-unique(which(is.na(HvN_degs_sig), arr.ind = T)[, 1]), ]
HvN_degs_sig$contrast <- "HvN"

LvN_degs <- read.csv(file.path(degs_dir, "LN.deg.csv"), stringsAsFactors = F)
LvN_degs <- LvN_degs[order(LvN_degs$pvalue), ]
LvN_degs_sig <- LvN_degs[LvN_degs$padj < 0.05, ]
LvN_degs_sig <- LvN_degs_sig[-unique(which(is.na(LvN_degs_sig), arr.ind = T)[, 1]), ]
LvN_degs_sig$contrast <- "LvN"

degs_sig <- rbind(HvL_degs_sig, HvN_degs_sig, LvN_degs_sig)
degs_sig_symbol <- degs_sig[, c("ENSEMBL", "SYMBOL", "contrast")]

#Enhancers
enhancers <- read.csv(file.path(parsed_enhancer_dir, "pcHiC_anno_GRCh38.csv"))
enhancers_oe <- enhancers[, c("oeChr", "oeStart", "oeEnd", "baitENS", "baitSYMBOL")]
enhancers_oe$regulator <- "enhancer"
colnames(enhancers_oe) <- c("chr", "start", "end", "ENS", "SYMBOL", "regulator")

#Promoters
promoters <- read.csv(file.path(parsed_promoter_dir, "GRCh38_promoters_blood.csv"), stringsAsFactors = F)
promoters_ENS <- promoters[, c("seqnames", "start", "end", "geneId", "SYMBOL")]
promoters_ENS$regulator <- "promoter"
colnames(promoters_ENS) <- c("chr", "start", "end", "ENS", "SYMBOL", "regulator")

enhancers_bait <- enhancers[, c("baitChr", "baitStart", "baitEnd", "baitENS", "baitSYMBOL")]
enhancers_bait$regulator <- "promoter"
colnames(enhancers_bait) <- c("chr", "start", "end", "ENS", "SYMBOL", "regulator")

promoters_total <- rbind(promoters_ENS, enhancers_bait)

#Regulators
regulators_gr <- makeGRangesFromDataFrame(rbind(promoters_ENS, enhancers_oe, enhancers_bait), keep.extra.columns = T)

#Find the regulators associated to the DEGs
deg_reg <- regulators_gr[regulators_gr$ENS %in% unique(degs_sig_symbol$ENS),]

#Find the genetic bodies of the regulators
require(GenomicFeatures)
txdb_dir <- file.path(data_dir, "Ensembl/Annotation")
grch38_txdb <- loadDb(file.path(txdb_dir, "Homo_sapiens.GRCh38.89.txdb.db"))

regulators_genes <- genes(grch38_txdb)[genes(grch38_txdb)$gene_id %in% regulators_gr$ENS, ]

ENStoSYMBOL <- unique(data.frame(mcols(regulators_gr))[, c(1, 2)])
rownames(ENStoSYMBOL) <- ENStoSYMBOL$ENS

regulators_genes$SYMBOL <- ENStoSYMBOL[regulators_genes$gene_id, "SYMBOL"]
colnames(mcols(regulators_genes)) <- c("ENS", "SYMBOL")
regulators_genes$regulator <- "body"

regulators_gr <- keepStandardChromosomes(c(regulators_gr, regulators_genes), pruning.mode = "coarse")
regulators_noSex_gr <- regulators_gr[!seqnames(regulators_gr) %in% c("X", "Y"),]
```

The first step in our analysis is to check out which RA-associated SNPs are located within our regions of interest. To this end, we will use the STOPGAP database. RA-associated SNPs will then be expanded by looking for SNPs in high LD (>0.8) after which the overlap will be returned. Note that STOPGAP is annotated using GRCh37, so the loci need to be uplifted to GRCh38.
```{r STOPGAP 2.5.1 data}
require(GenomicRanges)

#RA-GWAS
stopgap_dir <- file.path(data_dir, "STOPGAP2_5_1")

load(file.path(stopgap_dir, "gwas.RData"))
load(file.path(stopgap_dir, "ld.snps.r2.RData"))

head(gwas.data)
ra_gwas <- gwas.data[which(gwas.data$msh == "arthritis, rheumatoid"), ]

#LD
head(ld.snps.r2)
length(ra_gwas$rsid %in% ld.snps.r2$SNP.gwas)
ra_gwas_ld <- ld.snps.r2[ld.snps.r2$SNP.gwas %in% ra_gwas$snp_id, ]
ra_gwas_ld <- ra_gwas_ld[ra_gwas_ld$r2 > 0.8, ]
ra_gwas_ld$CHR.ld <- gsub("(.*)", "chr\\1", ra_gwas_ld$CHR.ld)
ra_gwas_ld_gr <- makeGRangesFromDataFrame(ra_gwas_ld, keep.extra.columns = T, seqnames.field = "CHR.ld", start.field = "POS.ld", end.field = "POS.ld")
names(ra_gwas_ld_gr) <- ra_gwas_ld_gr$SNP.ld

#LiftOver
require(rtracklayer)
hg19tohg38 <- import.chain(file.path(primary_analyses_dir, "data/analysis/WGBS/00correlations/hg19ToHg38.over.chain"))
ra_gwas_ld_gr <- unlist(liftOver(ra_gwas_ld_gr, hg19tohg38))
seqlevelsStyle(ra_gwas_ld_gr) <- "NCBI"

ra_gwas_ld_df <- data.frame(ra_gwas_ld_gr)[, c("SNP.ld", "seqnames", "start", "REF.ld", "ALT.ld", "SNP.gwas", "POS.gwas", "r2", "AF")]
colnames(ra_gwas_ld_df) <- c("SNP_ld", "chr_ld", "pos_ld", "ref_ld", "alt_ld", "SNP_gwas", "pos_gwas", "r2", "AF")
```

Can we find any SNPs located within the actual eQTMs?
```{r SNPS in DMRs}
gwas_output_dir <- file.path(output_dir, "06_GWAS_integration")
dir.create(gwas_output_dir)

SNP_overlapper <- function(dmrs_gr){
	dmr_gwas_overlap <- findOverlaps(ra_gwas_ld_gr, dmrs_gr) 
	dmrs_gwas_gr <- dmrs_gr[subjectHits(dmr_gwas_overlap),]

	mcols(dmrs_gwas_gr) <- cbind(mcols(dmrs_gwas_gr), ra_gwas_ld_df[queryHits(dmr_gwas_overlap),])

	return(dmrs_gwas_gr)
}

#aDMRs
admrs_gwas_gr <- SNP_overlapper(admrs_gr)
write.csv(data.frame(admrs_gwas_gr), file.path(gwas_output_dir, "aDMRs_internal_GWAS.csv"))

admrs_gwas_hr <- unique(data.frame(admrs_gwas_gr[,c("ENS", "SYMBOL", "REG", "rho", "rhoCI95_lower", "rhoCI95_upper", "SNP_ld", "ref_ld", "alt_ld")]))
admrs_gwas_hr <- aggregate(SNP_ld~., admrs_gwas_hr, FUN = toString)
write.csv(admrs_gwas_hr, file.path(gwas_output_dir, "aDMRs_internal_GWAS_HR.csv"))

admrs_gwas_gr_list <- split(admrs_gwas_gr, as.factor(admrs_gwas_gr))

#DaDMRs
dadmrs_gwas_gr <- SNP_overlapper(dadmrs_gr)
write.csv(data.frame(dadmrs_gwas_gr), file.path(gwas_output_dir, "DE_aDMRs_internal_GWAS.csv"))

#eQTMs
eqtms_gwas_gr <- SNP_overlapper(eqtms_gr)
write.csv(data.frame(eqtms_gwas_gr), file.path(gwas_output_dir, "eQTMs_internal_GWAS.csv"))

#DeQTMs
deqtms_gwas_gr <- SNP_overlapper(deqtms_gr)
write.csv(data.frame(deqtms_gwas_gr), file.path(gwas_output_dir, "DeQTMs_internal_GWAS.csv"))
```
Our data suggests that three SNPs were found in three DMRs, one of which was associated to two genes. While interesting, the genes associated to the DMRs were not considered to be differentially expressed. So instead, we broaden our search to look for SNPs in other regulatory regions of the eQTMs.

```{r SNPs in other regulatory elements of the genes associated to aDMRs and eQTMs }
##Find the RA-SNPs and the ones in LD that are located in the regulatory regions
snps_in_regulators <- findOverlaps(ra_gwas_ld_gr, regulators_gr)
regulators_with_snps <- regulators_gr[subjectHits(snps_in_regulators), ]
mcols(regulators_with_snps) <- cbind(mcols(regulators_with_snps), ra_gwas_ld_df[queryHits(snps_in_regulators), ])

dmr_snp_aggregator <- function(dmrs, snps){
	list(DMRs = dmrs, SNPs = snps)
}

#aDMRs
admrs_snps_gr <- regulators_with_snps[which(regulators_with_snps$ENS %in% admrs_gr$ENS),]
admrs_snps_gr <- makeGRangesFromDataFrame(unique(as.data.frame(admrs_snps_gr, row.names = NULL)[,-5]), keep.extra.columns = T)
admrs_snps_gr <- makeGRangesFromDataFrame(aggregate(regulator~., admrs_snps_gr, FUN = toString), keep.extra.columns = T)
admrs_dmrs_gr <- admrs_gr[which(admrs_gr$ENS %in% regulators_with_snps$ENS),]

##List object
admrs_snps_gr_list <- split(admrs_snps_gr, admrs_snps_gr$ENS)
admrs_dmrs_gr_list <- split(admrs_dmrs_gr, admrs_dmrs_gr$ENS)
admrs_snps_list <- mapply(dmr_snp_aggregator, admrs_dmrs_gr_list, admrs_snps_gr_list, SIMPLIFY = F)

##Dataframe object
genes_with_snps_and_dmrs <- do.call(rbind, lapply(admrs_snps_list, function(id){
	data.frame(DMR = paste(paste0(as.factor(id$DMRs), " (", id$DMRs$REG, ")"), collapse = ","),
		   SNP = paste(unique(id$SNPs$SNP_ld), collapse = ","),
		   ENS = unique(id$DMRs$SYMBOL))
}))

write.csv(data.frame(genes_with_snps_and_dmrs), file.path(gwas_output_dir, "aDMRs_SNPs.csv"))

##Dataframe object
#admrs_snps_df <- unique(data.frame(ENS = admrs_snps_gr$ENS, 
#				   SNP = admrs_snps_gr$SNP_ld, 
#				   REG = admrs_snps_gr$regulator,
#				   REF_ALLELE = admrs_snps_gr$ref_ld,
#				   ALT_ALLELE = admrs_snps_gr$alt_ld))
#admrs_dmrs_df <- unique(data.frame(SYMBOL = admrs_dmrs_gr$SYMBOL, 
#			    ENS = admrs_dmrs_gr$ENS, 
#			    DMR = paste0(as.factor(admrs_dmrs_gr)), 
#			    REG = admrs_dmrs_gr$REG))
#
#admrs_snps_summary <- merge(admrs_snps_df, admrs_dmrs_df, by.x = "ENS", by.y = "ENS")
#admrs_snps_summary <- admrs_snps_summary[,c(1,6,7,8,2,3,4,5)]
#colnames(admrs_snps_summary) <- c("ENS", "SYMBOL", "DMR", "DMR_REG", "SNP", "SNP_REG", "REF", "ALT") 
#
#write.csv(data.frame(admrs_snps_summary), file.path(gwas_output_dir, "aDMRs_SNPs.csv"))

#DaDMRs
dadmrs_snps_list <- admrs_snps_list[names(admrs_snps_list) %in% dadmrs_gr$ENS]
dadmrs_snps_summary <- admrs_snps_summary[rownames(admrs_snps_summary) %in% dadmrs_gr$ENS, ]

#eQTMs
eqtms_snps_list <- admrs_snps_list[names(admrs_snps_list) %in% eqtms_gr$ENS]
eqtms_snps_summary <- admrs_snps_summary[rownames(admrs_snps_summary) %in% eqtms_gr$ENS, ]

#DeQTMs
deqtms_snps_list <- admrs_snps_list[names(admrs_snps_list) %in% deqtms_gr$ENS]
deqtms_snps_summary <- admrs_snps_summary[rownames(admrs_snps_summary) %in% deqtms_gr$ENS, ]

```

```{r GWAS SNPs vs LD SNPs}
admrs_gwas <- admrs_snps_gr[admrs_snps_gr$SNP_ld == admrs_snps_gr$SNP_gwas,]
admrs_ld <- admrs_snps_gr[admrs_snps_gr$SNP_ld != admrs_snps_gr$SNP_gwas,]
```

```{r CD247}
CD247_SNPs <- admrs_snps_list$ENSG00000198821

CD247_ld <- unique(CD247_SNPs$SNPs$SNP_ld)
CD247_gwas <- unique(CD247_SNPs$SNPs$SNP_gwas)

ra_gwas[ra_gwas$snp_id %in% CD247_gwas,]

#Find out whether the SNPs locate within the DMRs
as.factor(CD247_SNPs$DMRs)
sort(unique(CD247_SNPs$SNPs$pos_ld))
```

```{r PHLDB1}
PHLDB1_SNPs <- admrs_snps_list$ENSG00000019144

PHLDB1_ld <- unique(PHLDB1_SNPs$SNPs$SNP_ld)
PHLDB1_gwas <- unique(PHLDB1_SNPs$SNPs$SNP_gwas)

ra_gwas[ra_gwas$snp_id %in% PHLDB1_gwas,]
```

```{r RPS6KA1}
RPS6KA1_SNPs <- admrs_snps_list$ENSG00000117676

RPS6KA1_ld <- unique(RPS6KA1_SNPs$SNPs$SNP_ld)
RPS6KA1_gwas <- unique(RPS6KA1_SNPs$SNPs$SNP_gwas)

ra_gwas[ra_gwas$snp_id %in% RPS6KA1_gwas,]
```

```{r STK17B}
STK17B_SNPs <- admrs_snps_list$ENSG00000081320

STK17B_ld <- unique(STK17B_SNPs$SNPs$SNP_ld)
STK17B_gwas <- unique(STK17B_SNPs$SNPs$SNP_gwas)

ra_gwas[ra_gwas$snp_id %in% STK17B_gwas,]
```

```{r PRKCB}
PRKCB_SNPs <- admrs_snps_list$ENSG00000166501

PRKCB_ld <- unique(PRKCB_SNPs$SNPs$SNP_ld)
PRKCB_gwas <- unique(PRKCB_SNPs$SNPs$SNP_gwas)

ra_gwas[ra_gwas$snp_id %in% PRKCB_gwas,]
```


```{r Visualization of the RA-SNPs in regulatory regions}
require(GenomicInteractions)
require(GenomicFeatures)
require(Gviz)

promoters_gr <- makeGRangesFromDataFrame(promoters_total, keep.extra.columns = T)

enhancers_bait <- makeGRangesFromDataFrame(df = enhancers, keep.extra.columns = T, seqnames.field = "baitChr", start.field = "baitStart", end.field = "baitEnd")
enhancers_oe <- makeGRangesFromDataFrame(df = enhancers, seqnames.field = "oeChr", start.field = "oeStart", end.field = "oeEnd")
enhancers_interact <- GenomicInteractions(enhancers_bait, enhancers_oe)

meth_dir <- file.path(primary_analyses_dir, "data/analysis/WGBS/07differentialMethylation")
meth <- readRDS(file.path(meth_dir, "meth.rds"))
seqlevelsStyle(meth) <- "NCBI"
factor_interest <- gsub("^.+\\.", "", colnames(mcols(meth)))

grch38_txdb <- keepStandardChromosomes(grch38_txdb)
grch38_gr <- genes(grch38_txdb)

require(biomaRt)
ENSBM <- useEnsembl(biomart = "ensembl", dataset = "hsapiens_gene_ensembl", version = 90)
ENS2Symbol <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"), filters = "ensembl_gene_id", values = names(grch38_gr), mart = ENSBM)
ENS2Symbol <- aggregate(hgnc_symbol~., ENS2Symbol, FUN = toString)
ENS2Symbol$hgnc_symbol <- gsub(", ", ";", ENS2Symbol$hgnc_symbol)
rownames(ENS2Symbol) <- ENS2Symbol$ensembl_gene_id
grch38_gr$SYMBOL <- ENS2Symbol[grch38_gr$gene_id, ]$hgnc_symbol

#ggplot colors
gg_color_hue <- function(n){
	hues <- seq(15, 375, length = n + 1)
	hcl(h = hues, l = 65, c = 100)[1:n]
}

#Gviz
require(Gviz)
aDMR_loc_core <- function(chr, start_region, end_region, measure = NULL, flanks = 1000, enh_int, dmrs_gr, snps_gr, prom_gr, gene_name, factor_interest, bm, build = "hg38", col_interaction = "red"){

	options(ucscChromosomeNames = F)
  if(build == "hg38") {gbuild <- "grch38"}

	#Plotting region
	eff_start <- start_region-flanks
	eff_end <- end_region+flanks

	gtrack <- GenomeAxisTrack()
	#itrack <- IdeogramTrack(genome = build, chromosome = chr)
	grtrack <- BiomartGeneRegionTrack(biomart = bm, genome = gbuild, chromosome = chr, start = eff_start, end = eff_end, name = "ENS", collapseTranscripts = "meta", transcriptAnnotation = "symbol")
	enhtrack <- InteractionTrack(enh_int, chromosome = chr, name = "Enh")
	displayPars(enhtrack) = list(anchor.height = 0.1,
				     col.interactions = col_interaction)
	if(!is.null(measure)){
		displayPars(enhtrack) = list(interaction.dimension = "height",
					     interaction.measure = measure)
	}

	promtrack <- AnnotationTrack(reduce(prom_gr), chromosome = chr, start = eff_start, end = eff_end, name = "Prom", stacking = "dense")
	methtrack <- AnnotationTrack(reduce(dmrs_gr), chromosome = chr, start = eff_start, end = eff_end, name = "DMR")
	snptrack <- AnnotationTrack(reduce(snps_gr), chromosome = chr, start = eff_start, end = eff_end, name = "RA-SNPs")
	
	tracklist <- list(gtrack, grtrack, enhtrack, promtrack, methtrack, snptrack) 
	
	plotTracks(trackList = tracklist, from = eff_start, to = eff_end, name = name, cex.title = 0.75, cex.axis = 0.75, cex.legend = 0.75, fontcolor = "black")
}

aDMR_viz <- function(ENS, dmr_snps_gr, factor_interest){
	#Find the plotting range to show the DMR, the gene and the regulatory region
	##Gene
	gene_features <- range(subsetByOverlaps(regulators_gr, dmr_snps_gr), ignore.strand = T)
	gene_coords <- grch38_gr[grch38_gr$gene_id == ENS,]
	gene_symbol <- ENS2Symbol[ENS, ]$hgnc_symbol
	plot_coords <- range(c(range(gene_features), range(gene_coords)), ignore.strand = T)

	#I know there is a function that flattens the GRanges object, I just cannot come up with it at the moment
	plot_chr <- as.character(seqnames(plot_coords))
	plot_start <- start(plot_coords)
	plot_end <- end(plot_coords)
	plot_flanks <- round(width(plot_coords)/5)

	#Filter enhancers_interact for the genes of interest
	enhancers_goi <- enhancers_interact[which(mcols(enhancers_interact)$anchor1.baitENS == ENS),]

	pdf(paste0(gene_symbol, ".pdf"), width = 8, height = 4)
	aDMR_loc_core(chr = plot_chr, 
		      start_region = plot_start, 
		      end_region = plot_end, 
		      enh_int = enhancers_goi, 
		      prom_gr = promoters_gr, 
		      bm = ENSBM, 
		      gene_name = gene_symbol, 
		      factor_interest = factor_interest, 
		      snps_gr = ra_gwas_ld_gr, 
		      dmrs_gr = admrs_gr, 
		      flanks = plot_flanks) 
	dev.off()
}

gwas_plot_dir <- file.path(gwas_output_dir, "genome_plots")
dir.create(gwas_plot_dir)

cur_dir <- getwd()

setwd(gwas_plot_dir)
#aDMR_viz(ENS = names(admrs_snps_list)[1], dmr_snps_gr = admrs_snps_list[[1]]$SNPs, factor_interest = factor_interest)

#Regulatory SNPs
deqtms_ens <- names(deqtms_snps_list)
for(i in length(deqtms_ens)){
	aDMR_viz(ENS = deqtms_ens[i], dmr_snps_gr = admrs_snps_list[[deqtms_ens[i]]]$SNPs, factor_interest = factor_interest)
}


```

```{r DMR with internal SNPs}
internal_SNP_dir <- file.path(gwas_plot_dir, "internal_SNPs")
dir.create(internal_SNP_dir)

aDMR_SNP_core <- function(meth, enh_int, dmrs_gr, snps_gr, prom_gr, gene_name, factor_interest, flanks = NULL, bm, build = "hg38", col_interaction = "red"){
	meth_gr <- subsetByOverlaps(meth, dmrs_gr)
	if(is.null(flanks)){
		flanks <- ceiling(width(dmrs_gr)/3)
	}
	dmrs_flanks_gr <- dmrs_gr + flanks

	options(ucscChromosomeNames = F)
  if(build == "hg38") {gbuild <- "grch38"}

	gtrack <- GenomeAxisTrack()
	#itrack <- IdeogramTrack(genome = build, chromosome = chr)
	grtrack <- BiomartGeneRegionTrack(biomart = bm, genome = gbuild, chromosome = as.character(seqnames(dmrs_flanks_gr)), start = start(dmrs_flanks_gr), end = end(dmrs_flanks_gr), name = "ENS", collapseTranscripts = "meta", transcriptAnnotation = "symbol")
	enhtrack <- InteractionTrack(enh_int, chromosome = as.character(seqnames(dmrs_flanks_gr)), name = "Enh")
	displayPars(enhtrack) = list(anchor.height = 0.1,
				     col.interactions = col_interaction)

	promtrack <- AnnotationTrack(reduce(prom_gr), chromosome = as.character(seqnames(dmrs_flanks_gr)), start = start(dmrs_flanks_gr), end = end(dmrs_flanks_gr), name = "Prom", stacking = "dense")
	#methtrack <- AnnotationTrack(reduce(dmrs_flanks_gr), chromosome = as.character(seqnames(dmrs_flanks_gr)), start = start(dmrs_flanks_gr), end = end(dmrs_flanks_gr), name = "DMR")
	methtrack <- DataTrack(meth_gr, groups = factor_interest, type = c("a", "p", "confint"), ylim = c(0,1), legend = T, col = gg_color_hue(length(levels(as.factor(factor_interest)))), name = "% Methylation")
	snptrack <- AnnotationTrack(reduce(snps_gr), chromosome = as.character(seqnames(dmrs_flanks_gr)), start = start(dmrs_flanks_gr), end = end(dmrs_flanks_gr), name = "RA-SNPs")
	
	tracklist <- list(gtrack, grtrack, enhtrack, promtrack, methtrack, snptrack) 
	htrack <- HighlightTrack(tracklist, chr = as.character(seqnames(dmrs_gr)), start = start(dmrs_gr), end = end(dmrs_gr), fill = NA)
	
	plotTracks(trackList = htrack, from = start(dmrs_flanks_gr), to = end(dmrs_flanks_gr), name = name, cex.title = 0.75, cex.axis = 0.75, cex.legend = 0.75, fontcolor = "black")
}

lapply(unique(admrs_gwas_gr), function(admr){
	pdf(file.path(internal_SNP_dir, gsub(":", "_", paste0(as.character(as.factor(admr)), ".pdf"))), width = 6, height = 6)
	aDMR_SNP_core(meth = meth, enh_int = enhancers_interact, dmrs_gr = admr, snps_gr = ra_gwas_ld_gr, prom_gr = promoters_gr, factor_interest = factor_interest, bm = ENSBM)
	dev.off()
})


#Internal SNPs
for(i in length(unique(admrs_gwas_gr))){
	aDMR_viz(unique(admrs_gwas_gr)[i,]$ENS, dmr_snps_gr = unique(admrs_gwas_gr)[i,], factor_interest = factor_interest)
}
setwd(cur_dir)

```

DMR 6:31286928-31287757 was found to display a SNP-like pattern across the entire DMR for two samples. within the DMR were 13 unique SNPs.
```{r putative mQTL}
putmqtl_snps <- admrs_gwas_gr[which(seqnames(admrs_gwas_gr) == 6 & start(admrs_gwas_gr) == 31286928),]$SNP_ld
putmqtl_snps_ld <- ra_gwas_ld[ra_gwas_ld$SNP.ld %in% putmqtl_snps,]


#rs146805812
soi_gwas <- unique(admrs_gwas_gr[which(admrs_gwas_gr$SNP_ld == "rs146805812"),]$SNP_gwas)
ra_gwas[ra_gwas$rsid %in% soi_gwas,]

ra_gwas_ld[ra_gwas_ld$SNP.gwas %in% soi_gwas,]

```



The second step is the overrepresentation analysis of the RA-associated SNPs. We will utilize the Ensembl Variant database (v89), which contains common variants from dbSNP, 1000 genomes and HAPMAP. 
```{r Overrepresentation analysis - Tentative}
require("VariantAnnotation")
ensvar_dir <- "../../../../data/ensembl_variation/"
ensvar <- readVcf(file.path(ensvar_dir, "Homo_sapiens.vcf"), "GRCh38")

```
